# Basin Characteristics

The basin characteristics (aka covariates) dataset provides a variety of landscape metrics associated with the NECD catchments. These covariates were computed to support the EcoSHEDS stream temperature and brook trout occupancy model. The available variables include:

```{r cov-list, message=FALSE}
covariates <- read_csv("../covariates/EcoSHEDS NECD Covariates List.csv", col_types = cols(.default = col_character())) |> 
  select(group = Group, name = Name, units = `Final Units`, description = `Data Layer`, data_source = `Data Source`)
covariates |> 
  mutate(group = factor(group, levels = c("Land Cover", "Human Impacts", "Topography", "Geology", "Climate"))) |> 
  arrange(group, name) |> 
  knitr::kable(align = "lllll", col.names = c("Group", "Variable", "Units", "Description", "Source"))
```


**Documentation:** The following spreadsheet contains a more detailed list of the covariates. 

```{r cov-copy}
invisible(file.copy("../covariates/EcoSHEDS NECD Covariates List.xlsx", "./downloads/EcoSHEDS NECD Covariates List.xlsx"))
```

- [EcoSHEDS NECD Covariates List.xlsx](./downloads/EcoSHEDS NECD Covariates List.xlsx)

**Downloads:** the covariate datasets are available for download in pre-staged HUC2 regions (corresponding to the [Catchment Delineation] shapefiles)

- [covariates_01.zip](`r paste0(base_url, "covariates/covariates_01.zip")`): Region 01
- [covariates_02.zip](`r paste0(base_url, "covariates/covariates_02.zip")`): Region 02
- [covariates_03.zip](`r paste0(base_url, "covariates/covariates_03.zip")`): Region 03
- [covariates_04.zip](`r paste0(base_url, "covariates/covariates_04.zip")`): Region 04
- [covariates_05.zip](`r paste0(base_url, "covariates/covariates_05.zip")`): Region 05
- [covariates_06.zip](`r paste0(base_url, "covariates/covariates_06.zip")`): Region 06

Each zip file contains four files:

- `Catchments##.csv`: values computed over entire catchment polygons
- `zonalStatsForDB_riparianBufferDetailed50ft_##.csv`: values computed within 50 ft riparian buffer
- `zonalStatsForDB_riparianBufferDetailed100ft_##.csv`: values computed within 100 ft riparian buffer
- `zonalStatsForDB_riparianBufferDetailed200ft_##.csv`: values computed within 200 ft riparian buffer

Each csv file contains the following columns:

```{r cov-columns}
tibble::tribble(
  ~name, ~description,
  "featureid", "catchment ID (joins to FEATUREID attribute in catchments shapefile)",
  "variable", "variable name",
  "value", "value",
  "zone", "`local` = value computed within catchment polygon<br>`upstream` = value computed within this and all upstream catchments",
  "riparian_distance_ft", "riparian buffer distance (ft)<br>if `NA` then value computed over entire catchment, otherwise value is computed only within this distance from the flowlines"
) |> 
  knitr::kable(align = "ll", col.names = c("Column", "Description"))
```

